---
title: "mix6"
author: "Allison Bray"
date: "November 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load in the data
```{r}
train = read.table('train.csv', sep = ',', header = 1, row.names = NULL)
test = read.table('test.csv', sep = '-', header = 1, row.names = NULL)
library(mice)
library(dplyr)

```

remove unused columns 
```{r}
train[, "DAY"] = NULL
train[, "HOUR"] = NULL
train[, "MINUTE"] = NULL
train[, "USAF"] = NULL
train[, "YEAR"] = NULL
train[, "MONTH"] = NULL
```


make a training set dataframe
```{r}
trainDF = train
```

make it a matrix
```{r}
train_mat = as.matrix(trainDF)
head(train_mat)
```
mice
```{r}
imp <- mice(train_mat, method = "norm.predict", m = 1) # Impute data
data_sto <- complete(imp) # Store data
mice_mat = as.matrix(data_sto)
```


inputting mice into any nan where there are six or more nan values in a row 
```{r}
count = 0
for(i in 1:(dim(train_mat)[1])){
  
  if(is.nan(train_mat[i,1])){
    count = count + 1
  }
  else if(count > 5){
    x = i-count
    y = i-1
      for(j in x:y){
       train_mat[j,1] = mice_mat[j,1]
       count = 0
    }
  }
  else{
    count = 0
  }
}

count = 0
for(i in 1:(dim(train_mat)[1])){
  
  if(is.nan(train_mat[i,2])){
    count = count + 1
  }
  else if(count > 5){
    x = i-count
    y = i-1
      for(j in x:y){
       train_mat[j,2] = mice_mat[j,2]
       count = 0
    }
  }
  else{
    count = 0
  }
}

count = 0
for(i in 1:(dim(train_mat)[1])){
  
  if(is.nan(train_mat[i,3])){
    count = count + 1
  }
  else if(count > 5){
    x = i-count
    y = i-1
      for(j in x:y){
       train_mat[j,3] = mice_mat[j,3]
       count = 0
    }
  }
  else{
    count = 0
  }
}

count = 0
for(i in 1:(dim(train_mat)[1])){
  
  if(is.nan(train_mat[i,4])){
    count = count + 1
  }
  else if(count > 5){
    x = i-count
    y = i-1
      for(j in x:y){
       train_mat[j,4] = mice_mat[j,4]
       count = 0
    }
  }
  else{
    count = 0
  }
}

count = 0
for(i in 1:(dim(train_mat)[1])){
  
  if(is.nan(train_mat[i,5])){
    count = count + 1
  }
  else if(count > 5){
    x = i-count
    y = i-1
      for(j in x:y){
       train_mat[j,5] = mice_mat[j,5]
       count = 0
    }
  }
  else{
    count = 0
  }
}
```


##############################

## fill with push forward 
```{r}
#train_mat = as.matrix(trainDF)
pushf = na.locf(train_mat, na.rm = TRUE, fromLast = TRUE)
```



## fill with push backward 
```{r}
pushb = na.locf(train_mat, na.rm = TRUE, fromLast = FALSE)
#train_mat = as.matrix(trainDF)

pushb[1,1] = -0.131349547
```


### avg pushf and pushb results
```{r}
pushfb = (pushf + pushb)/2

head(pushfb)


```


```{r}
for(i in 1:(dim(train_mat)[1])){
   for(j in 1:5){
    if(is.na(train_mat[i,j])){
      train_mat[i,j] = pushfb[i,j]
    }
  }
}
```


```{r}
n.test = dim(test)[1]

test_mat = as.matrix(test)
result_mat_mix = test[,2]
```

using test.csv as ID key
retrieve the new corresponding imputed values from pushfb matrix
```{r}
i = 1
for(i in 1:n.test){
  test_rindex = as.integer(test_mat[i,1])
  test_cindex = as.integer(test_mat[i,2])
  result_mat_mix[i] = train_mat[test_rindex,test_cindex]
}

head(result_mat_mix)

test_ = read.table('test.csv', header = 1, row.names = NULL)
```
create df for output, col 1 ID, col 2 imputed values specified by ID key
```{r}
final_result_mix = data.frame(ID = as.vector(test_), 
                          Value = as.vector(result_mat_mix))
```
Output data to csv
```{r}
write.table(final_result_mix,
            file = 'result_mix6.csv',
            sep = ',',
            quote = F,
            row.names = FALSE
)
```


